Solving Collective Spin Hamiltonians with Decoherence

### 2018-08-25 soc_squeezing/cumulant_expansions.tex
added explenation of operator sorting in TAT-zy case

\newcommand{\Z}{\mathbb{Z}}

\newcommand{\bmu}{{\bar\mu}}

@@

Previous work has shown how to realize various spin-squeezing
Hamiltonians with ultracold atomic systems.  While it is
straightforward to simulate the dynamics generated by these
Hamiltonians in the context of a closed quantum system, here we try to
develop an understanding of their the squeezing behavior in the
presence of decoherence.  In particular, we will be considering
dynamics induced by the collective one-axis twisting (OAT), two-axis
twisting (TAT), and transverse field (TVF) Hamiltonians of the form
\begin{align}
  H_{\t{OAT}} = \chi S_z^2,
  &&
  H_{\t{TAT}}^{z,y} = \chi \p{S_z^2 - S_y^2},
  &&
  H_{\t{TVF}} = \chi S_z^2 - \Omega S_x,
\end{align}
for spins which decay in an uncorrelated manner at a rate $\gamma$.
The evolution of any Heisenburg operator $\O$ with a Hamiltonian $H$
is given by
\begin{align}
  \f{d}{dt} \O
  = i\sp{H,\O}_- + \gamma\D_-\p{\O},
  &&
  \D_-\p{\O} \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
    - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+},
  \label{eq:EOM}
\end{align}
where $\sp{A,B}_\pm\equiv AB\pm BA$ and $j$ indexes an individual
spin.  Our measure of squeezing is
\begin{align}
  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $N$ is the total number of spins, $\v S$ is the collective spin
vector, and the minimization is performed over all unit vectors
$\uv n$ in the plane orthogonal to the mean spin vector $\bk{\v S}$.
This squeezing parameter is entirely determined by components
$\bk{S_\alpha}$ of the mean spin vector in addition to collective
spin-spin correlators of the form $\bk{S_\alpha S_\beta}$; our task is
therefore to compute these expectation values in all scenarios of
interest.

@@

The case of the one-axis twisting is simplest to consider, as we can
write the Hamiltonian in the form
\begin{align}
  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_j^z \sigma_k^z
  = \f12 \chi \sum_{j<k} \sigma_j^z \sigma_k^z + \f14 N \chi,
\end{align}
which admits an exact solution previously worked out by Foss-Feig et
al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
expectation values with respect to the initial state $\ket{X}$ which
has all spins pointing in $x$, in terms of $\mu,\nu\in\set{+1,-1}$ we
find that
\begin{align}
  \bk{S_+}
  &= \sum_j\bk{\sigma_j^+}
  = \f{N}{2} e^{-\gamma t/2} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
  \bk{S_\mu S_z}
  &= \f12\sum_j\bk{\sigma_j^\mu\sigma_j^z}
  + \f12\sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^z} \\
  &= -\f{\mu}{2}\bk{S_\mu} + \f14 N \p{N-1} e^{-\gamma t/2}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{S_\mu S_\nu}
  &= \sum_j \bk{\sigma_j^\mu \sigma_j^\nu}
  + \sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^\nu} \\
  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_z}}
  + \f14 N \p{N-1} e^{-\gamma t}
  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \sp{\cos\p{s_\zeta t}+\f12\gamma t~\t{sinc}\p{s_\zeta t}},
  &&
  \Psi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \p{is_\zeta-\gamma/2}t~ \t{sinc}\p{s_\zeta t},
\end{align}
and $s_\zeta \equiv \zeta + i2\gamma$.  The last ingredients we need
to compute spin squeezing at any time are $\bk{S_z}$ and $\bk{S_z^2}$.
As these operators commute with the OAT Hamiltonian, their evolution
is governed entirely by decoherence, such that
\begin{align}
  \f{d}{dt}\p{S_z^n}
  = \gamma \D_-\p{S_z^n}
  = \gamma \p{S + S_z} \sum_{k=0}^{n-1}
  \p{-1}^{n-k} { n \choose k } S_z^k,
\end{align}
where $S\equiv N/2$ and $\D_-\p{S_z^n}$ is worked out in Appendix
\ref{sec:decoherence}.  In particular,
\begin{align}
  \f{d}{dt} S_z = -\gamma \p{S + S_z},
  &&
  \f{d}{dt}\p{S_z^2} = \gamma \p{S + S_z} \p{1 - 2S_z}.
\end{align}
The initial conditions $\bk{S_z}_{t=0}=0$ and $\bk{S_z^2}_{t=0}=S/2$
then imply
\begin{align}
  \bk{S_z} = S\p{e^{-\gamma t}-1},
  &&
  \bk{S_z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_z}^2.
  \label{eq:Sz_OAT}
\end{align}
The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
$\xi^2$ in \eqref{eq:squeezing} at any time throughout evolution of
the initial state $\ket{X}$ under $H_{\t{OAT}}$.

@@

The equations of motion for $S_\mu$, $S_\mu S_z$, and $S_\mu S_\nu$
(i.e.~where $\mu,\nu\in\set{+1,-1}$) under the OAT Hamiltonian
$H_{\t{OAT}}=\chi S_z^2$ are determined by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
  = i\chi \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n},
\end{align}
where we can use identity \eqref{eq:comm_z_mu} to simplify
\begin{align}
  \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  = \sp{S_z^2, S_+^\ell S_-^m}_- S_z^n
  = S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n,
  \label{eq:S_z^2_comm}
\end{align}
and the effect of decoherence through $\D_-\p{S_+^\ell S_-^m S_z^n}$ is
worked out in Appendix \ref{sec:decoherence}.  In order to compute
squeezing of the OAT Hamiltonian in the presence of single-spin decay,
we therefore need to solve the systems of equations defined by
\begin{align}
  \f{d}{dt} \p{S_\mu S_z^n}
  = i\chi S_\mu \p{1 + 2\mu S_z} S_z^n
  + \gamma S_\mu \sp{\p{S - \delta_{\mu,-1} + S_z} \sum_{k=0}^{n-1}
    \p{-1}^{n-k} { n \choose k } S_z^k - \f12 S_z^n}
  \label{eq:evolution_OAT_single}
\end{align}
and
\begin{align}
  \f{d}{dt} \p{S_\mu S_\nu S_z^n}
  &= i\chi S_\mu S_\nu \p{\sp{\mu+\nu}^2 + 2 \sp{\mu+\nu} S_z} S_z^n
  \tag*{} \\
  &\quad + \gamma S_\mu S_\nu
  \sp{\p{S - \delta_{\mu,-1} - \delta_{\nu,-1} + S_z} \sum_{k=0}^{n-1}
    \p{-1}^{n-k} { n \choose k } S_z^k - S_z^n},
  \eqref{eq:evolution_OAT_pair}
\end{align}
for $\mu\ge\nu$.  The initial conditions for these systems of
equations can be determined from the expansions
\begin{align}
  S_\mu = \sum_{m=-S}^S \sqrt{\p{S-\mu m}\p{S+\mu m+1}} \op{m+\mu}{m},
  &&
  S_z^n = \sum_{m=-S}^S m^n \op{m},
\end{align}
and the initial state
\begin{align}
  \ket{X} = 2^{-N/2} \sum_{m=-S}^S { N \choose S+m }^{1/2} \ket{m},
\end{align}
from which we can determine
\begin{align}
  \bk{S_\mu S_z^n}_{t=0}
  &= 2^{-N} \sum_{m=-S}^S m^n
  \sp{{ N \choose S+m } { N \choose S+m+\mu }
    \p{S-\mu m}\p{S+\mu m+1}}^{1/2} \\
  &= 2^{-N} \sum_{m=-S}^S m^n { N \choose S+m } \abs{m - \mu S},
\end{align}
and
\begin{multline}
  \bk{S_\mu S_\nu S_z^n}_{t=0}
  = 2^{-N} \sum_{m=-S}^S m^n
  \sp{{ N \choose S+m } { N \choose S+m+\mu+\nu }}^{1/2} \\
  \times \sp{\p{S-\nu\sp{m+\nu}}\p{S+\mu\sp{m+\nu}+1}
    \p{S-\nu m}\p{S+\nu m+1}}^{1/2}.
\end{multline}

@@

All collective spin Hamiltonians can be written in the form
\begin{align}
  H = \sum_{p,q,r} h_{pqr} S_+^p S_-^q S_z^r.
  \label{eq:H_general}
\end{align}
The evolution of a general operator $S_+^\ell S_-^m S_z^n$ under a
Hamiltonian of the form in \eqref{eq:H_general} is given by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_-^q S_z^r, S_+^\ell S_-^m S_z^n}_-
  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n},
\end{align}
where the commutator is given in Appendix \ref{sec:comm_general}, and
the decoherence operator in Appendix \ref{sec:decoherence}.

@@

The transverse-field (TVF) twisting Hamiltonian with unit twisting
strength is
\begin{align}
  H_{\t{TVF}}
  = S_z^2 - \Omega S_x
  = S_z^2 - \f{\Omega}{2} S_+ - \f{\Omega}{2} S_-.
\end{align}
Unlike in the case of OAT, we now cannot divide all operators of the
form $S_+^\ell S_-^m S_z^n$ into independent sectors of constant
imbalances $\ell,m$.  We therefore need to keep track of all operators
in one vector $\v V_{\t{TVF}}$, which in turn requires imposing an
ordering on all operators of the form $S_+^\ell S_-^m S_z^n$.  Such an
ordering is provided by the one-to-one map
$p:\mathbb{N}_0^3\to\mathbb{N}_0$, where $\mathbb{N}_0$ denotes the
set of non-negative integers, given by
\begin{align}
  v\p{\ell,m,n} = B\p{T_{\ell mn}} + R\p{T_{\ell mn},\ell} + m,
  \label{eq:vec_map}
\end{align}
where
\begin{align}
  T_{\ell mn} \equiv \ell + m + n,
  &&
  B\p{T} \equiv \f16 T \p{T+1} \p{T+2},
  &&
  R\p{T,\ell} \equiv \f12 \ell \p{2 T - \ell + 3}.
\end{align}
This map is designed to sort integers $\p{\ell,m,n}$ first by their
sum $T_{\ell mn}$ (i.e.~the total weight of $S_+^\ell S_-^m S_z^n$),
and then in turn by the values of $\ell$, $m$, and $n$, such that
$v\p{\ell,m,n}<v\p{\ell',m',n'}$ if
\begin{enumerate*}[label=(\roman*)]
\item $T_{\ell mn}<T_{\ell'm'n'}$,
\item $T_{\ell mn}=T_{\ell'm'n'}$ and $\ell<\ell'$,
\item $\p{T_{\ell mn},\ell}=\p{T_{\ell'm'n'},\ell'}$ and $m<m'$, or
\item $\p{T_{\ell mn},\ell,m}=\p{T_{\ell'm'n'},\ell',m'}$ and $n<n'$.
\end{enumerate*}
Fixing a maximal operator weight $M$, the vector keeping track of all
operators is then
\begin{align}
  \v V_{\t{TVF}}
  \equiv \sum_{\ell+m+n\le M} S_+^\ell S_-^m S_z^n \uv x_{v\p{\ell,m,n}}.
\end{align}

@@

The two-axis twisting Hamiltonian one axis oriented along the
decoherence axis is
\begin{align}
  H_{\t{TAT}}^{z,y}
  = S_z^2 - S_y^2
  = S_z^2 + \f14 S_+^2 + \f14 S_-^2 - \f12 S_+ S_- + \f12 S_z.
\end{align}
While upper and lower imbalances are not conserved by this
Hamiltonians, the parity of the net imbalance, i.e.~$\ell-m~\t{mod}~2$
for an operator $S_+^\ell S_-^m S_z^n$, is conserved.  There are
therefore two sectors of operators which we can evolve independently
to save computational overhead.

>>>>

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}

\newcommand{\bmu}{{\bar\mu}}
\newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}

@@

Previous work has shown how to realize various spin-squeezing
Hamiltonians with ultracold atomic systems.  While it is
straightforward to simulate the dynamics generated by these
Hamiltonians in the context of a closed quantum system, here we try to
develop an understanding of their the squeezing behavior in the
presence of decoherence.  In particular, we will be considering
dynamics induced by the collective one-axis twisting (OAT), two-axis
twisting (TAT), and transverse field (TVF) Hamiltonians of the form
\begin{align}
  H_{\t{OAT}} = \chi S_z^2,
  &&
  H_{\t{TAT}}^{\p{z,y}} = \chi \p{S_z^2 - S_y^2},
  &&
  H_{\t{TVF}} = \chi S_z^2 + \Omega S_x,
\end{align}
for spins which decay in an uncorrelated manner at a rate $\gamma$.
The evolution of any Heisenburg operator $\O$ with a Hamiltonian $H$
is given by
\begin{align}
  \f{d}{dt} \O
  = i\sp{H,\O}_- + \gamma\D_-\p{\O},
  &&
  \D_-\p{\O} \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
    - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+},
  \label{eq:EOM}
\end{align}
where $\sp{A,B}_\pm\equiv AB\pm BA$ and $j$ indexes an individual
spin.  Our measure of squeezing is
\begin{align}
  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $N$ is the total number of spins, $\v S$ is the collective spin
vector, and the minimization is performed over all unit vectors
$\uv n$ in the plane orthogonal to the mean spin vector $\bk{\v S}$.
This squeezing parameter is entirely determined by components
$\bk{S_\alpha}$ of the mean spin vector in addition to collective
spin-spin correlators of the form $\bk{S_\alpha S_\beta}$; our task is
therefore to compute these expectation values in all scenarios of
interest.

@@

The case of the one-axis twisting is simplest to consider, as we can
write the Hamiltonian in the form
\begin{align}
  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_j^z \sigma_k^z
  = \f12 \chi \sum_{j<k} \sigma_j^z \sigma_k^z + \f14 N \chi,
\end{align}
which admits an exact solution previously worked out by Foss-Feig et
al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
expectation values with respect to the initial state $\ket{X}$
satisfying $S_x\ket{X}=S\ket{X}$, in terms of $\mu,\nu\in\set{+1,-1}$
we find that
\begin{align}
  \bk{S_+}
  &= \sum_j\bk{\sigma_j^+}
  = \f{N}{2} e^{-\gamma t/2} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
  \bk{S_\mu S_z}
  &= \f12\sum_j\bk{\sigma_j^\mu\sigma_j^z}
  + \f12\sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^z} \\
  &= -\f{\mu}{2}\bk{S_\mu} + \f14 N \p{N-1} e^{-\gamma t/2}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{S_\mu S_\nu}
  &= \sum_j \bk{\sigma_j^\mu \sigma_j^\nu}
  + \sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^\nu} \\
  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_z}}
  + \f14 N \p{N-1} e^{-\gamma t}
  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \sp{\cos\p{s_\zeta t}+\f12\gamma t~\t{sinc}\p{s_\zeta t}},
  &&
  \Psi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \p{is_\zeta-\gamma/2}t~ \t{sinc}\p{s_\zeta t},
\end{align}
and $s_\zeta \equiv \zeta + i2\gamma$.  The last ingredients we need
to compute spin squeezing at any time are $\bk{S_z}$ and $\bk{S_z^2}$.
As these operators commute with the OAT Hamiltonian, their evolution
is governed entirely by decoherence, such that
\begin{align}
  \f{d}{dt}\p{S_z^n}
  = \gamma \D_-\p{S_z^n}
  = \gamma \p{S + S_z} \sum_{k=0}^{n-1}
  \p{-1}^{n-k} { n \choose k } S_z^k,
\end{align}
where $S\equiv N/2$ and $\D_-\p{S_z^n}$ is worked out in Appendix
\ref{sec:decoherence}.  In particular,
\begin{align}
  \f{d}{dt} S_z = -\gamma \p{S + S_z},
  &&
  \f{d}{dt}\p{S_z^2} = \gamma \p{S + S_z} \p{1 - 2S_z}.
\end{align}
The initial conditions $\bk{S_z}_{t=0}=0$ and $\bk{S_z^2}_{t=0}=S/2$
then imply
\begin{align}
  \bk{S_z} = S\p{e^{-\gamma t}-1},
  &&
  \bk{S_z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_z}^2.
  \label{eq:Sz_OAT}
\end{align}
The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
$\xi^2$ in \eqref{eq:squeezing} at any time throughout evolution of
the initial state $\ket{X}$ under $H_{\t{OAT}}$.

@@

The equations of motion for $S_\mu$, $S_\mu S_z$, and $S_\mu S_\nu$
(i.e.~where $\mu,\nu\in\set{+1,-1}$) under the OAT Hamiltonian
$H_{\t{OAT}}=\chi S_z^2$ are determined by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
  = i\chi \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n},
\end{align}
where we can use identity \eqref{eq:comm_z_mu} to simplify
\begin{align}
  \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  = \sp{S_z^2, S_+^\ell S_-^m}_- S_z^n
  = S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n,
  \label{eq:S_z^2_comm}
\end{align}
and the effect of decoherence through $\D_-\p{S_+^\ell S_-^m S_z^n}$ is
worked out in Appendix \ref{sec:decoherence}.  In order to compute
squeezing of the OAT Hamiltonian in the presence of single-spin decay,
we therefore need to solve the systems of equations defined by
\begin{align}
  \f{d}{dt} \p{S_\mu S_z^n}
  = i\chi S_\mu \p{1 + 2\mu S_z} S_z^n
  + \gamma S_\mu \sp{\p{S - \delta_{\mu,-1} + S_z} \sum_{k=0}^{n-1}
    \p{-1}^{n-k} { n \choose k } S_z^k - \f12 S_z^n}
  \label{eq:evolution_OAT_single}
\end{align}
and
\begin{multline}
  \f{d}{dt} \p{S_\mu S_\nu S_z^n}
  = i\chi S_\mu S_\nu \p{\sp{\mu+\nu}^2 + 2 \sp{\mu+\nu} S_z} S_z^n \\
  + \gamma S_\mu S_\nu
  \sp{\p{S - \delta_{\mu,-1} - \delta_{\nu,-1} + S_z} \sum_{k=0}^{n-1}
    \p{-1}^{n-k} { n \choose k } S_z^k - S_z^n},
  \label{eq:evolution_OAT_pair}
\end{multline}
for $\mu\ge\nu$.  The initial conditions for these systems of
equations can be determined from the expansions
\begin{align}
  S_z^n = \sum_{m=-S}^S m^n \op{m},
  &&
  S_\mu = \sum_{m=-S}^S g_\mu\p{m} \op{m+\mu}{m},
\end{align}
with
\begin{align}
  g_\mu\p{S,m} \equiv \sqrt{\p{S-\mu m}\p{S+\mu m+1}},
  \label{eq:transverse_elem}
\end{align}
and the initial state
\begin{align}
  \ket{X} = 2^{-N/2} \sum_{m=-S}^S { N \choose S+m }^{1/2} \ket{m},
\end{align}
from which we can determine
\begin{align}
  \bk{S_\mu S_z^n}_{t=0}
  = 2^{-N} \sum_{m=-S}^S m^n { N \choose S+m } \abs{m - \mu S},
\end{align}
and
\begin{align}
  \bk{S_\mu S_\nu S_z^n}_{t=0}
  = 2^{-N} \sum_{m=-S+\delta_{\mu,-1}+\delta_{\nu,-1}}
  ^{S-\delta_{\mu,1}-\delta_{\nu,1}}
  m^n \sp{{ N \choose S+m+\mu+\nu } { N \choose S+m }}^{1/2}
  g_\mu\p{S,m+\nu} g_\nu\p{S,m}.
\end{align}

@@

All collective spin Hamiltonians can be written in the form
\begin{align}
  H = \sum_{p,q,r} h_{pqr} S_+^p S_-^q S_z^r.
  \label{eq:H_general}
\end{align}
The evolution of a general operator $S_+^\ell S_-^m S_z^n$ under a
Hamiltonian of the form in \eqref{eq:H_general} is given by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_-^q S_z^r, S_+^\ell S_-^m S_z^n}_-
  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n},
\end{align}
where the commutator is given in Appendix \ref{sec:comm_general}, and
the decoherence operator in Appendix \ref{sec:decoherence}.  We will
be considering evolution of the initial states
$\ket{X}:S_x\ket{X}=S\ket{X}$ and $\ket{-Z}:S_z\ket{-Z}=-S\ket{-Z}$,
for which initial conditions are given by
\begin{multline}
  \bk{X|S_+^\ell S_-^m S_z^n|X}
  =  2^{-N} \sum_{k=-S+m}^{S-\ell} k^n
  \sp{{ N \choose S+k+\ell-m } { N \choose S+k }}^{1/2} \\
  \times \prod_{p=0}^{\ell-1} g_+\p{S,k-m+p}
  \prod_{q=0}^{m-1} g_-\p{S,k-q}
\end{multline}
and
\begin{multline}
  \bk{-Z|S_+^\ell S_-^m S_z^n|-Z}
  = \p{-S}^n 2^{-N} \sum_{k=-S+m}^{S-\ell}
  \sp{{ N \choose S+k+\ell-m } { N \choose S+k }}^{1/2} \\
  \times \prod_{p=0}^{\ell-1} g_+\p{S,k-m+p}
  \prod_{q=0}^{m-1} g_-\p{S,k-q}.
\end{multline}

@@

The transverse-field (TVF) twisting Hamiltonian with unit twisting
strength is
\begin{align}
  H_{\t{TVF}}
  = S_z^2 + \Omega S_x
  = S_z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-.
\end{align}
Unlike in the case of OAT, we now cannot divide all operators of the
form $S_+^\ell S_-^m S_z^n$ into independent sectors of constant
imbalances $\ell,m$.  We therefore need to keep track of all operators
in one vector $\v V_{\t{TVF}}$, which in turn requires imposing an
ordering on all operators of the form $S_+^\ell S_-^m S_z^n$.  Such an
ordering is provided by the one-to-one map $u:\N_0^3\to\N_0$, where$
\N_0$ denotes the set of non-negative integers, given by
\begin{align}
  u\p{\ell,m,n} = B\p{T_{\ell mn}} + R\p{T_{\ell mn},\ell} + m,
  \label{eq:vec_map}
\end{align}
where $T_{\ell mn}\equiv\ell+m+n$ for shorthand and
\begin{align}
  B\p{T}
  \equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 : T_{\ell mn}<T}}
  = \f16 T \p{T+1} \p{T+2},
\end{align}
\begin{align}
  R\p{T,L}
  \equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 : T_{\ell mn}=T , \ell<L}}
  = \f12 \ell \p{2 T - \ell + 3}.
\end{align}
In words, $u\p{\ell,m,n}$ sorts the collection non-negative integers
$\p{\ell,m,n}$ first by their sum $T_{\ell mn}$, and then in turn by
the values of $\ell$ and $m$.  Fixing a maximal operator weight $M$,
the vector keeping track of all operators is then
\begin{align}
  \v V_{\t{TVF}}
  \equiv \sum_{\ell+m+n\le M} S_+^\ell S_-^m S_z^n \uv x_{u\p{\ell,m,n}}.
\end{align}

@@

The two-axis twisting Hamiltonian one axis oriented along the
decoherence axis is
\begin{align}
  H_{\t{TAT}}^{\p{z,y}}
  = S_z^2 - S_y^2
  = S_z^2 + \f14 S_+^2 + \f14 S_-^2 - \f12 S_+ S_- + \f12 S_z.
\end{align}
While upper and lower imbalances are not conserved by this
Hamiltonians, the parity of the net imbalance, i.e.~$\ell-m~\t{mod}~2$
for an operator $S_+^\ell S_-^m S_z^n$, is conserved.  There are
therefore two sectors of operators which we can evolve independently
to save computational overhead.  To keep track of these operators in
one-dimensional vectors, we define the onto maps $v_p:\N_0^3\to\N_0$
for $p\in\set{0,1}$ given by
\begin{align}
  v_p\p{\ell,m,n} = B_p\p{T_{\ell mn}} + R_p\p{T_{\ell mn},n} + m,
\end{align}
where
\begin{align}
  B_p\p{T}
  &\equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 :
      T_{\ell mn}<T, \ell-m=p~\t{mod}~2}} \\
  &= \f12 B\p{T} + \f14 \p{T+1} \p{T~\t{mod}~2} \p{1-2p},
\end{align}
\begin{align}
  R_p\p{T,N}
  &\equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 :
      T_{\ell mn}=T, \ell-m=p~\t{mod}~2, n < N }} \\
  &= \floor{\f{n}{2}}
  \p{T - \floor{\f{n}{2}} + 2 - \p{T+p~\t{mod}~2}}.
\end{align}
The map $v_p\p{\ell,m,n}$ sorts the collection non-negative integers
$\p{\ell,m,n}$ with $\ell-m=p~\t{mod}~2$ first by their sum
$T_{\ell mn}$, and then in turn by the values of $n$ and $m$.  Fixing
a maximum operator weight $M$, the vectors keeping track of all
operators are then
\begin{align}
  \v V_{\t{TAT}}^{\p{z,y,p}}
  \equiv \sum_{\substack{\ell,m,n\le M\\\ell-m=p~\t{mod}~2}}
  S_+^\ell S_-^m S_z^n \uv x_{v_p\p{\ell,m,n}}.
\end{align}
